---
title: "class15"
author: "Shabnam Hamdard"
date: "11/14/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import count and metadata

```{r}
library(DESeq2)

metaFile <- "data/GSE37704_metadata.csv"
countFile <- "data/GSE37704_featurecounts.csv"

# Import metadata and take a peak
colData = read.csv(metaFile, row.names=1)
head(colData)
```

```{r}
countData = read.csv(countFile, row.names=1)
head(countData)
```

The "length" column in countData is going to cause problems with the required matching to the metadata file, so let's remove it here.

```{r}
countData <- as.matrix(countData[,-1])
head(countData)
```

Double check that the colnames in countData match the id values in the colData metadata file.

```{r}
colnames(countData)
```

```{r}
rownames(colData)
```

We can use the all() function to check if all entries in a vector are TRUE.
```{r}
all( colnames(countData) == rownames(colData) )
```

```{r}
# Test how the all() function works!
all( c(T,F,T) )
```

# Remove zero count genes

We want to remove genes that have 0 count values in alll experiments (i.e. rows that have 0 across all cols).

```{r}
countData = countData[rowSums(countData) != 0,]
head(countData)
```

## DESeq2 Analysis

```{r}
library(DESeq2)

# Setup the object with our data in the way DESeq wants it

dds = DESeqDataSetFromMatrix(countData=countData,
                             colData=colData,
                             design=~condition)
#Run the analysis
dds = DESeq(dds)
```

```{r}
res = results(dds)
res
```

```{r}
plot( res$log2FoldChange, -log(res$padj) )
```

# Add some color!
```{r}
mycols <- rep("gray", length(res$padj))
mycols[ abs(res$log2FoldChange) > 2] <- "blue"
mycols[ res$padj > 0.005 ] <- "gray"
plot( res$log2FoldChange, -log(res$padj), col=mycols )
```

```{r}
library(org.Hs.eg.db)

columns(org.Hs.eg.db)
```

```{r}
res$symbol = mapIds(org.Hs.eg.db,
                    keys= row.names(countData), 
                    keytype="ENSEMBL",
                    column="SYMBOL",
                    multiVals="first")

res
```

```{r}
res$entrez = mapIds(org.Hs.eg.db,
                    keys=row.names(countData),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")
```

```{r}
res$name =   mapIds(org.Hs.eg.db,
                    keys=row.names(res),
                    keytype="ENSEMBL",
                    column="GENENAME",
                    multiVals="first")
```

## Pathway analysis

```{r}
library(gage)
library(gageData)
```

``` {r}
data(kegg.sets.hs)
data(sigmet.idx.hs)

# Focus on signaling and metabolic pathways only
kegg.sets.hs = kegg.sets.hs[sigmet.idx.hs]

# Examine the first 3 pathways
head(kegg.sets.hs, 3)
```

```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```

```{r}
keggres = gage(foldchanges, gsets=kegg.sets.hs)
```

```{r}
attributes(keggres)
```

```{r}
head(keggres$less)
```

```{r}
library(pathview)

pathview(gene.data=foldchanges, pathway.id="hsa04110")
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa04110", kegg.native=FALSE)
```

```{r}
## Focus on top 5 upregulated pathways here for demo purposes only
keggrespathways <- rownames(keggres$greater)[1:5]

# Extract the 8 character long IDs part of each string
keggresids = substr(keggrespathways, start=1, stop=8)
keggresids
```

```{r}
pathview(gene.data=foldchanges, pathway.id=keggresids, species="hsa")
```

```{r}
krp <- rownames(keggres$less)[1:5]

krids = substr(krp, start=1, stop=8)
krids
```

```{r}
pathview(gene.data=foldchanges, pathway.id=krids, species="hsa")
```

# Section 3

```{r}
data(go.sets.hs)
data(go.subs.hs)

# Focus on Biological Process subset of GO
gobpsets = go.sets.hs[go.subs.hs$BP]

gobpres = gage(foldchanges, gsets=gobpsets, same.dir=TRUE)

lapply(gobpres, head)
```

