---
title: "class09"
author: "Shabnam Hamdard"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Input


```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```

Note that the `id` and `diagnosis` columns will not be used for most of the following steps.

We have `r nrow(wisc.df)` samples in this dataset.

```{r}
nrow(wisc.df)
```

How many benign (not cancerous) and malignant (cancerous) samples do we have in the dataset?

```{r}
table(wisc.df$diagnosis)
```

```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[3:32])

# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id

head(wisc.data)
```

Store the diagnosis for reference in the future as a separate vector!

```{r}
diagnosis <- wisc.df$diagnosis
```

## Questions

- Q1. How many observations are in this dataset?
```{r}
nrow(wisc.df)
```

- Q2. How many of the observations have a malignant diagnosis?
```{r}
table(wisc.df$diagnosis)
```

- Q3. How many varialbles/features in the data are suffixed with mean?
```{r}
colnames(wisc.df)
```

```{r}
grep("_mean", colnames(wisc.df), val= TRUE)
```

I can use `length()` to count how many matches we have.

```{r}
length(grep("_mean", colnames(wisc.df), val= TRUE))
```

```{r}
colMeans(wisc.data)
```

```{r}
apply(wisc.data, 2, sd)
```



```{r}
round( colMeans(wisc.data), 3)
```

```{r}
round( apply(wisc.data, 2, sd), 3)
```

These values look very different so I will use `scale=TRUE` when I run PCA.

```{r}
# Perform PCA on wisc.data
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```





Let's make a plot of PC1 vs. PC2

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis)
```

Color by cancer vs. non-cancer
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis)
```

```{r}
biplot(wisc.pr)
```

```{r}
pc.var <- wisc.pr$sdev^2
head(pc.var)
```

```{r}
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wc.hclust <- hclust(data.dist)
plot(wc.hclust)
abline(h = 20, col="red", lty=2)
```

```{r}
wisc.hclust.clusters <- cutree(wc.hclust, k=4)
table(wisc.hclust.clusters, diagnosis)
```

```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:7]), "ward.D2")
plot(wisc.pr.hclust)
grps <- cutree(wisc.pr.hclust, k=2)
```

```{r}
table(grps)
table(grps, diagnosis)
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
```

```{r}
g <- as.factor(grps)
levels(g)

g <- relevel(g,2)
levels(g)

plot(wisc.pr$x[,1:2], col=g)
```

```{r}
url <- "https://tinyurl.com/new-samples-CSV"

new <- read.csv(url)

npc <- predict(wisc.pr, newdata=new)

npc
```

Plot these new samples on our PC1 vs PC2 plot...

```{r}
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

- Q17. Which of these new patients should we prioritize for follow up based on your results?

The patients from Group 1.