---
title: "class08"
author: "Shabnam Hamdard"
date: "10/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)
```

Use the kmeans() function setting k to 2 and nstart=20
```{r}
k <- kmeans(x, centers = 2, nstart = 20)
```

Inspect/print the results
```{r}
k
```

Q. How many points are in each cluster? -> 30 points
Q. What ‘component’ of your result object details
 - cluster size?
 - cluster assignment/membership?
 - cluster center?
```{r}
k$size
```

```{r}
k$cluster
```

```{r}
k$centers
```

Plot x colored by the kmeans cluster assignment and
      add cluster centers as blue points

```{r}
plot(x, col=k$cluster) # 1 - black; 2 - red
points(k$centers, col="blue", pch=15)
```


## Hierarchical Clustering
The `hclust()` function requires a distance matrix as input. You can get this from the `dist()` function

```{r}
# First we need to calculate point (dis)similarity
# as the Euclidean distance between observations
dist_matrix <- dist(x)

# The hclust() function returns a hierarchical
# clustering model
hc <- hclust(dist_matrix)

# the print method is not so useful here
hc

```

```{r}
plot(hc)
abline(h=6, col="red")
grps <- cutree(hc, h=6) # Cut by height h
```

```{r}
cutree(hc, k=2)
```

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")

# Step 2. Plot the data without clustering
plot(x)

# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )

plot(x, col=col)
```

```{r}
# clustering
hc <- hclust( dist(x) )

# draw tree
plot(hc)
abline(h=3, col="tomato")

# cut the tree
grps <- cutree(hc, k=3)
grps
```

```{r}
plot(x, col=grps)
```

```{r}
table(grps)
```

Cross-tabulate (i.e. compare our clustering result with the known answer).

```{r}
table(grps, col)
```

## Principal Component Analysis (PCA)

```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV", row.names=1)

head(mydata)
```

How many genes are in this dataset?
```{r}
dim(mydata)
nrow(mydata)
```

Let's do PCA with the **prcomp()** function.
```{r}
pca <- prcomp(t(mydata), scale=TRUE)

plot(pca$x[,1], pca$x[,2])

pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

pca.var.per
```

```{r}
barplot(pca.var.per, main="Scree Plot", xlab="Prinicpal Component", ylab="Percent Variation")
```

```{r}
plot(pca$x[,1], pca$x[,2],
     col=c("red", "red", "red", "red", "red",
           "blue", "blue", "blue", "blue", "blue"))
```

```{r}
x <- read.csv("UK_foods.csv", row.names=1)
head(x)
```

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

```{r}
barplot(as.matrix(x), col=rainbow(nrow(x)))
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```

```{r}
pca <- prcomp( t(x) )
summary(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x), col=c("orange", "red", "blue", "green"))
```

