---
title: "class14"
author: "Shabnam Hamdard"
date: "11/12/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Title

```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <- read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
```

```{r}
head(metadata)
```

How many genes do we have in this dataset?
```{r}
nrow(counts)
```

```{r}
nrow(metadata)
```

We want to know if there is a difference in expression values for control (non-drug) vs. treated (drug added cell lines).

First step is to find which experiments were the control experiments and then get the average values across all control experiments. Then we will do the same for drug treated.

```{r}
control <- metadata[metadata[,"dex"]=="control",]

control.mean <- rowSums( counts[ ,control$id] )/length(control$id)

names(control.mean) <- counts$ensgene
```

```{r}
treated <- metadata[metadata[, "dex"]=="treated",]

treated.mean <- rowSums( counts[ ,treated$id] )/length(treated$id)

names(treated.mean) <- counts$ensgene
```

We will combine our meancount data for bookkeeping purposes.

```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

Plot control vs treated!

```{r}
plot(meancounts$control.mean, meancounts$treated.mean, log="xy")
```

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])

head(meancounts)
```

**which** is used to show which values in control and treatment are equal to 0.

**unique** is used to simplify the values returned by which.

The rows that return a value of zero are then removed from the dataset.

```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
sum(up.ind)
```

```{r}
sum(down.ind)
```

```{r}
head(mycounts[up.ind,])
```

## DESeq2 analysis

```{r}
library(DESeq2)

dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

```{r}
# Dull all black plot
plot( res$log2FoldChange, -log(res$padj))
abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.05), col="gray", lty=2)
```

```{r}
# Add some color please
mycols <- rep("gray", nrow(res))

# Makes points with +2 or -2 fold change blue
mycols[abs(res$log2FoldChange) > 2] <- "blue"

# Make points below p-value cutoff gray
mycols[ (res$padj > 0.05) ] <- "gray"

plot( res$log2FoldChange, -log(res$padj), col=mycols)
abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.05), col="gray", lty=2)
```

## Save our results for next day...

```{r}
write.csv(res, file="expression_results.csv")
```

