---
title: "class07"
author: "Shabnam Hamdard"
date: "10/22/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Functions Revisited

Source my functions from last day

```{r}
source("http://tinyurl.com/rescale-R")
```


```{r}
rescale(1:10)
```

```{r}
rescale( c(1, 10, 5, NA, 6))
```

```{r}
# rescale2(c(1:10, "pizza"))
```

```{r}
# rescale(c(1:10, "pizza"))
```

## A new function called both_na()
Write a function to find where there are NA elements in two input vectors.


First make some simple input where I know the answer.
```{r}
x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

Looked online and found the **is.na** function.
```{r}
is.na(x)
```

And the **which()** function tells me where the true values are.
```{r}
which(is.na(x))
```

```{r}
which(is.na(y))
```

```{r}
is.na(x)
is.na(y)
```

The AND function requires two input TRUE to give a TRUE output.
```{r}
is.na(x) & is.na(y)
```

Taking the **sum()** of TRUE FALSE vector will tell me how many TRUE elements I have. This is my working snippet!
```{r}
sum(is.na(x) & is.na(y))
```


```{r}
sum( c(TRUE, TRUE, FALSE, TRUE) )
```


Now turn it into a function.


```{r}
both_na <- function(x, y) {
  sum( is.na(x) & is.na(y) )
}
```

```{r}
both_na(c(1, NA, 3, NA, 4), c(3, NA, 4, 2, NA))
```

Test Function

```{r}
x <- c(NA, NA, NA)
y1 <- c(1, NA, NA)
y2 <- c(1, NA, NA, NA)
```

```{r}
both_na(x, y2)
```

```{r}
x2 <- c(NA, NA)
both_na(x2, y2)
```


```{r}
plot(1:10, col=c("red", "blue", "green"))
```

```{r}
x3 <- c(1, NA, NA)
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x3, y3)
```

```{r}
x3 <- c(1, NA, NA, 1, NA, NA, 1)
y3 <- c(1, NA, NA, NA, NA, NA, NA)
```

```{r}
length(x)
length(y3)
```

Add a check for when inputs x and y are not the same length.
```{r}
both_na2 <- function(x, y) {
  if(length(x) != length(y)) {
    stop("Input x and y are not of equal length!")
  }
  sum( is.na(x) & is.na(y) )
}
```

```{r}
# both_na2(x, y3)
```


```{r}
# student 1
s1 <- c(100, 100, 100, 100, 100, 100, 100, 90)

# student 2
s2 <- c(100, NA, 90, 90, 90, 90, 97, 80)

mean(s1[ -which.min(s1) ])
```

```{r}
which.min(s2)
s2[ -which.min(s2) ]
mean(s2 [- which.min(s2) ], na.rm = TRUE)
```

We have our working code. Now, we can turn it into a function.

```{r}
grade <- function(x) {
  if (any(is.na(x)) ) {
    warning("Student is missing a homework!")
  }
  mean( x[-which.min(x) ], na.rm=TRUE)
}
```

```{r}
grade(s1)
grade(s2)
```

```{r}
url <- "https://tinyurl.com/gradeinput"
hw <- read.csv(url, row.names=1)
```

```{r}
apply(hw, 1, grade)
```


```{r}
grades <- function(x) {
  ordered_list <- sort(x, decreasing = TRUE, na.last = TRUE)
  new_grades <- head(ordered_list, -1)
  print(paste("The average is", mean(new_grades)))
  return(new_grades)
}

grades(s1)
grades(s2)
```

```{r}
grade2 <- function(x) {
  if (any(is.na(x)) ) {
    warning("Student is missing a homework!")
  }
  sum(mean( x[-which.min(x) ], na.rm=TRUE))/(length(x)-1)
}
```

```{r}
s3 <- c(100, NA, NA, NA, NA)
grade2(s3)
```

